---
title: "Reproducible Analysis: The Two-Wing admixed structure of environmental microbial communities"
author: "Yushi Tang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r library, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)       # for generating plots
library(latex2exp)     # for plot text latex
library(cowplot)       # for merging plots
library(gridExtra)     # for griding plots
library(grid)          # for griding plots
library(stringr)       # for uppercase first letter
```

# Preface

This document provides reproducible research records for our manuscript about the Two-Wing admixed structure of environmental microbial communities. We provide step-by-step instructions for main results in both the original article and the supplemental materials.

# Main Figures 

## Figure 1. Observed abundance distribution

### The over-dispersion pattern of microbial communities

```{r mean_var_plot_hz, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
load('../output/HZ/hz_mean_var.RData')
dat_plt$Region <- factor(dat_plt$Region, levels=c('Bay','ERA JX','ERA SY'))
ggplot(dat_plt, aes(x=miseqmean, y=miseqvar)) + 
    geom_point(alpha=0.36, size=1.5) +
    xlab(TeX('Abundance Mean ($\\mu$)')) +
    ylab(TeX('Abundance Variance ($v$)')) +
    xlim(0,3000) +
    ylim(0,4e7) +
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.title.x = element_text(size = 10),
          axis.title.y = element_text(size = 10),
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7)) +
    theme(aspect.ratio=1) +
    facet_grid(~Region)
```


```{r mean_var_plot_ao, eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
load('../output/AO_ASV/ao_mean_var.RData')
dat_plt$Industry <- factor(dat_plt$Industry, levels=c('Dye','Pharmaceutical','Pesticide'))
dat_plt$Process <- factor(dat_plt$Process, levels=c('Influent','Anoxic','Oxic','Effluent'))
ggplot(dat_plt, aes(x=miseqmean, y=miseqvar)) + 
    geom_point(alpha=0.36, size=1.5) +
    xlab(TeX('Abundance Mean ($\\mu$)')) +
    ylab(TeX('Abundance Variance ($v$)')) +
    xlim(0,1000) +
    ylim(0,4e6) +
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.title.x = element_text(size = 10),
          axis.title.y = element_text(size = 10),
          axis.text.x = element_text(size = 7),
          axis.text.y = element_text(size = 7)) +
    theme(aspect.ratio=1) +
    facet_grid(Industry~Process)
```

## Figure 2. The *Two-Wing* admixed structure

### Before boundary refining (Figure 2A)

```{r two_wing_prerefine, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
param_plt <- data.frame()
# load AO parameter trio
vec_process <- c('influent','anoxic','oxic','effluent')
for(item in vec_process){
  file_in <- paste0('../output/AO_ASV/param_trio_asv_',item,'.RData')
  load(file_in)
  param_trio$ID <- rownames(param_trio)
  # extract poisson distributed microbes
  param_poisson <- param_trio[param_trio$k==Inf,]
  # extract gamma-poisson distributed microbes
  param_gpm <- param_trio[param_trio$k!=Inf,]
  # extract zero-inflated microbes
  param_zeroinf <- param_trio[param_trio$pi0!=0,]
  # annotate procedure
  param_gpm$Procedure <- str_to_title(item)
  param_plt <- rbind(param_plt, param_gpm)
}

# load HZ parameter trio
vec_process <- c('bay','era')
for(item in vec_process){
  file_in <- paste0('../output/HZ/param_trio_',item,'.RData')
  load(file_in)
  param_trio$ID <- rownames(param_trio)
  # extract poisson distributed microbes
  param_poisson <- param_trio[param_trio$k==Inf,]
  # extract gamma-poisson distributed microbes
  param_gpm <- param_trio[param_trio$k!=Inf,]
  # extract zero-inflated microbes
  param_zeroinf <- param_trio[param_trio$pi0!=0,]
  # annotate procedure
  param_gpm$Procedure <- str_to_title(item)
  param_plt <- rbind(param_plt, param_gpm)
}

rownames(param_plt) <- c(1:nrow(param_plt))
param_plt$Procedure <- ifelse(param_plt$Procedure=='Era','ERA',param_plt$Procedure)

param_plt$Procedure <- factor(param_plt$Procedure,
                              levels=c('Bay','Influent','Oxic', 
                                       'ERA','Anoxic','Effluent'))

param_plt$mu_up <- 4.457417
param_plt$mu_up <- ifelse(param_plt$Procedure=='ERA', 
                          4.881282, param_plt$mu_up)
param_plt$mu_up <- ifelse(param_plt$Procedure=='Influent', 
                          7.142101, param_plt$mu_up)
param_plt$mu_up <- ifelse(param_plt$Procedure=='Anoxic', 
                          5.209871, param_plt$mu_up)
param_plt$mu_up <- ifelse(param_plt$Procedure=='Oxic', 
                          4.299989, param_plt$mu_up)
param_plt$mu_up <- ifelse(param_plt$Procedure=='Effluent', 
                          4.405847, param_plt$mu_up)


param_plt$k_low <- 10.452344
param_plt$k_low <- ifelse(param_plt$Procedure=='ERA', 
                          10.452344, param_plt$k_low)
param_plt$k_low <- ifelse(param_plt$Procedure=='Influent', 
                          8.844581, param_plt$k_low)
param_plt$k_low <- ifelse(param_plt$Procedure=='Anoxic', 
                          8.673801, param_plt$k_low)
param_plt$k_low <- ifelse(param_plt$Procedure=='Oxic', 
                          8.673801, param_plt$k_low)
param_plt$k_low <- ifelse(param_plt$Procedure=='Effluent', 
                          8.812485, param_plt$k_low)


ggplot() +
  geom_point(data=param_plt, 
             aes(x=log(k), y=log(mu), color=pi0),alpha=0.6,size=0.24) +
  geom_vline(data=param_plt, aes(xintercept = mu_up), 
             linetype="dotdash", color = "black", size=0.12) +
  geom_vline(data=param_plt, aes(xintercept = k_low), 
             linetype="dotdash", color = "black", size=0.12) +
  ylab(TeX('\\log{(\\hat{\\mu})}')) +
  xlab(TeX('\\log{(\\hat{k})}')) +
  ggtitle(TeX('')) +
  xlim(-5,20) +
  ylim(-5,10) +
  theme_bw() +
  theme(aspect.ratio=1) +
  scale_color_continuous(type = "viridis") +
  facet_wrap(Procedure~., scale='free', nrow=2) +
  labs(color=TeX('$\\hat{\\pi}_0$')) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(size = 9, vjust = 0.5, hjust = 0),
        axis.text.y = element_text(size=9),
        strip.text = element_text(size = 12),
        legend.title = element_text(color = "black", size = 12),
        legend.text = element_text(color = "black", size = 9))

```

### After boundary refining (Figure 2B)

```{r two_wing_postrefine, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# load parameters mle
vec_process <- c('influent','anoxic','oxic','effluent')
param_plt <- data.frame()
for(item in vec_process){
  load(paste0('../output/AO_ASV/param_trio_asv_',item,'.RData'))
  load(paste0('../output/AO_ASV/mic_od_',item,'.RData'))
  param_trio$ID <- rownames(param_trio)
  # extract gamma-poisson distributed microbes
  param_gpm <- param_trio[param_trio$k!=Inf,]
  # annotate procedure
  param_gpm$Procedure <- str_to_title(item)
  # annotate taxa category
  load(paste0('../output/AO_ASV/mic_id_mu_k_',item,'.RData'))
  param_gpm$Wing <- ifelse(param_gpm$ID %in% id_mu, 'mu', 'Other')
  param_gpm$Wing <- ifelse(param_gpm$ID %in% id_k,  'k', param_gpm$Wing)
  param_plt <- rbind(param_plt, param_gpm)

}

vec_process <- c('bay','era')
for(item in vec_process){
  load(paste0('../output/HZ/param_trio_',item,'.RData'))
  load(paste0('../output/HZ/mic_od_',item,'.RData'))
  param_trio$ID <- rownames(param_trio)
  # extract gamma-poisson distributed microbes
  param_gpm <- param_trio[param_trio$k!=Inf,]
  # annotate procedure
  param_gpm$Procedure <- str_to_title(item)
  # annotate taxa category
  load(paste0('../output/HZ/mic_id_mu_k_',item,'.RData'))
  param_gpm$Wing <- ifelse(param_gpm$ID %in% id_mu, 'mu', 'Other')
  param_gpm$Wing <- ifelse(param_gpm$ID %in% id_k, 'k', param_gpm$Wing)
  
  param_plt <- rbind(param_plt, param_gpm)

}

rownames(param_plt) <- c(1:nrow(param_plt))
param_plt$Procedure <- ifelse(param_plt$Procedure=='Era','ERA',param_plt$Procedure)

param_plt$Procedure <- factor(param_plt$Procedure,
                              levels=c('Bay','Influent','Oxic', 
                                       'ERA','Anoxic','Effluent'))
param_plt$Wing <- factor(param_plt$Wing,
                             levels=c('mu','k', 'Other'))

ggplot() +
  geom_point(data=param_plt, 
             aes(x=log(k), y=log(mu), color=Wing), alpha=0.36,size=0.6) +
  ylab(TeX('\\log{(\\hat{\\mu})}')) +
  xlab(TeX('\\log{(\\hat{k})}')) +
  ggtitle(TeX('')) +
  xlim(-5,20) +
  ylim(-5,10) +
  theme_bw() +
  theme(aspect.ratio=1) +
  facet_wrap(Procedure~., scale='free') +
  scale_colour_manual(values = c('#984ea3','#4daf4a'),
                      labels = unname(TeX(c('$\\mu$','$k$')))) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(size = 9, vjust = 0.5, hjust = 0),
        axis.text.y = element_text(size=9),
        strip.text = element_text(size = 12),
        legend.title = element_text(color = "black", size = 10),
        legend.text = element_text(color = "black", size = 9)) +
  guides(color = guide_legend(override.aes = list(size = 1.8)))

```

```{r SI, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
sessionInfo()
```

