---
title: "Reproducible Analysis: The Two-Wing admixed structure of environmental microbial communities"
author: "Yushi Tang"
date: "`r Sys.Date()`"
output: pdf_document
---

# Preface

This document provides reproducible research records for our manuscript about the *Two-Wing* admixed structure of environmental microbial communities. We provide step-by-step instructions for main results in both the original article and the supplemental materials.

# Data and Directories

We would suggest to set up the same directories as the *Two-Wing* paper repository at <https://github.com/YushiFT/two-wing-mic>. Open this markdown file under `your/path/to/docs`. Unzip the two data sets under their original directory, which includes

* HZ data: <https://github.com/YushiFT/two-wing-mic/tree/main/data/HZ/hz_asv_2018.txt.zip>

* AO data: <https://github.com/YushiFT/two-wing-mic/tree/main/data/AO/ao_asv.txt.zip>

# Packages

## Main package

Our analysis relies on the `R` package `PMCosm` (Version 0.1.5) developed by our lab. Software handbook and quick instructions about using `PMCosm` are available at <https://github.com/YushiFT/PMCosm>. We first install and load the main package `PMCosm`.

```{r install_main, echo=TRUE, eval=FALSE, warning=FALSE, message=FALSE}
# change to path/to/docs
# setwd(your/path/to/docs)
install.packages("devtools")
devtools::install_github("YushiFT/PMCosm")
```

```{r library_main, echo=TRUE, eval=TRUE, warnin=FALSE, message=FALSE}
library(PMCosm)
```

## Other packages

```{r library_other, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)       # for generating plots
library(latex2exp)     # for plot text latex
library(cowplot)       # for merging plots
library(gridExtra)     # for griding plots
library(grid)          # for griding plots
library(stringr)       # for uppercase first letter
library(ggh4x)         # for grid plot with free axis
library(vegan)         # for example data in pcoa analysis
```

# Community Structure Inference

## Estimate MLE trio

This step can take several hours. Cluster computing recommended. 

```{r mle_trio, echo=TRUE, eval=FALSE, warning=FALSE, message=FALSE}
# hz data
mic <- read.table(file='../data/HZ/hz_asv_2018.txt',
                  header=TRUE,row.names=1) # 51441 taxa in total
# construct column id
bay_loci <- c('HB1.1','HB1.2','HB1.3','HB2.1','HB2.2','HB2.3',
              'HB3.1','HB3.2','HB3.3','HB4.1','HB4.2','HB4.3',
              'HB5.1','HB5.2','HB5.3','HB6.1','HB6.2','HB6.3',
              'HB7.1','HB7.2','HB7.3','HB8.1','HB8.2','HB8.3',
              'HB9.1','HB9.2','HB9.3','HB10.1','HB10.2','HB10.3')
era_sy_loci <- c('SY1.1','SY1.2','SY1.3','SY2.1','SY2.2','SY2.3',
                 'SY3.1','SY3.2','SY3.3','SY4.1','SY4.2','SY4.3',
                 'SY5.1','SY5.2','SY5.3','SY6.1','SY6.2','SY6.3')
era_jx_loci <- c('JX1.1','JX1.2','JX1.3','JX2.1','JX2.2','JX2.3',
                 'JX3.1','JX3.2','JX3.3','JX4.1','JX4.2','JX4.3',
                 'JX5.1','JX5.2','JX5.3','JX6.1','JX6.2','JX6.3')

# for hz bay
mic_bay <- mic[,bay_loci] # 51441 taxa
# filter records with zero counts across all sample sites
mic_bay <- mic_bay[rowSums(mic_bay)>0,] # 24383 taxa
# save log records to ../output/HZ
sink(file='../output/HZ/mle_trio_trace_bay.txt')
# estimate mle trio
param_trio <- calc_mle_trio(mic_bay, n_sample=10, replicates=3)
# end log file
sink()
save(param_trio, file='../output/HZ/param_trio_bay.RData')

# for hz era
mic_era <- mic[,c(era_sy_loci,era_jx_loci)] # 51441 taxa
# filter records with zero counts across all sample sites
mic_era <- mic_era[rowSums(mic_era)>0,] # 33606 taxa
# save log records to ../output/HZ
sink(file='../output/HZ/mle_trio_trace_era.txt')
# estimate mle trio
param_trio <- calc_mle_trio(mic_era, n_sample=12, replicates=3)
# end log file
sink()
save(param_trio, file='../output/HZ/param_trio_era.RData')


# ao data
# import the whole data set
mic <- read.table('../data/AO/ao_asv.txt')
# industrial factory names
id_dye <- c('LS','SF','CZ','BA','YF')
id_med <- c('GB','YTSW','XHC','ZC')
id_pes <- c('YT','YN')
lis_ind <- list(id_dye, id_med, id_pes)
names(lis_ind) <- c('Dye','Medicine','Pesticide')
# ao procedures id
id_inf <- c('Inf1','Inf2','Inf3')
id_axi <- c('Ax1','Ax2','Ax3')
id_oxi <- c('Ox1','Ox2','Ox3')
id_eff <- c('Eff1','Eff2','Eff3')
lis_pro <- list(id_inf, id_axi, id_oxi, id_eff)
names(lis_pro) <- c('Influent','Anoxic','Oxic','Effluent')

# ao influent
region <- 1
id_col <- c()
# extract samples in target procedure/region
for(i in 1:length(lis_ind)){
    id_ind <- lis_ind[[i]]
    id_pro <- lis_pro[[region]]
    id_col <- c(id_col,
                paste0(rep(id_ind, each=length(id_pro)),'_',rep(id_pro, length(id_ind))))
}
mic_sub <- mic[,id_col] # 111981 taxa
mic_sub <- mic_sub[rowSums(mic_sub)>0,] # 16763 taxa
# save log records to ../output/AO
sink(file='../output/AO/mle_trio_trace_influent.txt')
# estimate mle trio
param_trio <- calc_mle_trio(mic_sub, n_sample=11, replicates=3)
# end log file
sink()
save(param_trio, file='../output/AO/param_trio_influent.RData')


# ao anoxic
region <- 2
id_col <- c()
# extract samples in target procedure/region
for(i in 1:length(lis_ind)){
    id_ind <- lis_ind[[i]]
    id_pro <- lis_pro[[region]]
    id_col <- c(id_col,
                paste0(rep(id_ind, each=length(id_pro)),'_',rep(id_pro, length(id_ind))))
}
mic_sub <- mic[,id_col] # 111981 taxa
mic_sub <- mic_sub[rowSums(mic_sub)>0,] # 31922 taxa
# save log records to ../output/AO
sink(file='../output/AO/mle_trio_trace_anoxic.txt')
# estimate mle trio
param_trio <- calc_mle_trio(mic_sub, n_sample=11, replicates=3)
# end log file
sink()
save(param_trio, file='../output/AO/param_trio_anoxic.RData')


# ao oxic
region <- 3
id_col <- c()
# extract samples in target procedure/region
for(i in 1:length(lis_ind)){
    id_ind <- lis_ind[[i]]
    id_pro <- lis_pro[[region]]
    id_col <- c(id_col,
                paste0(rep(id_ind, each=length(id_pro)),'_',rep(id_pro, length(id_ind))))
}
mic_sub <- mic[,id_col] # 111981 taxa
mic_sub <- mic_sub[rowSums(mic_sub)>0,] # 36071 taxa
# save log records to ../output/AO
sink(file='../output/AO/mle_trio_trace_oxic.txt')
# estimate mle trio
param_trio <- calc_mle_trio(mic_sub, n_sample=11, replicates=3)
# end log file
sink()
save(param_trio, file='../output/AO/param_trio_oxic.RData')


# ao effluent
region <- 4
id_col <- c()
# extract samples in target procedure/region
for(i in 1:length(lis_ind)){
    id_ind <- lis_ind[[i]]
    id_pro <- lis_pro[[region]]
    id_col <- c(id_col,
                paste0(rep(id_ind, each=length(id_pro)),'_',rep(id_pro, length(id_ind))))
}
mic_sub <- mic[,id_col] # 111981 taxa
mic_sub <- mic_sub[rowSums(mic_sub)>0,] #  38330 taxa
# save log records to ../output/AO
sink(file='../output/AO/mle_trio_trace_effluent.txt')
# estimate mle trio
param_trio <- calc_mle_trio(mic_sub, n_sample=11, replicates=3)
# end log file
sink()
save(param_trio, file='../output/AO/param_trio_effluent.RData')

```

## Visualize the *Two-Wing* admixed structure by MLE trio estimates

### HZ data

The left panel below is HZ Bay, right panel is HZ ERA. 

```{r mle_trio_hz, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/HZ/param_trio_bay.RData')
g1 <- plot_trio(param_trio, zoom_in=TRUE)

load('../output/HZ/param_trio_era.RData')
g2 <- plot_trio(param_trio, zoom_in=TRUE)

plot_grid(g1, g2, nrow=1, align='h')
```

### AO data

From left top to right bottom, the order is *Influent*, *Anoxic*, *Oxic*, and *Effluent*.

```{r mle_trio_ao, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/AO/param_trio_influent.RData')
g1 <- plot_trio(param_trio, zoom_in=TRUE)

load('../output/AO/param_trio_anoxic.RData')
g2 <- plot_trio(param_trio, zoom_in=TRUE)

load('../output/AO/param_trio_oxic.RData')
g3 <- plot_trio(param_trio, zoom_in=TRUE)

load('../output/AO/param_trio_effluent.RData')
g4 <- plot_trio(param_trio, zoom_in=TRUE)

plot_grid(g1, g2, nrow=1, align='h')

plot_grid(g3, g4, nrow=1, align='h')
```

# Dispersal Vanguards and Laggards

## Clarify the model-driven testable boundary

### HZ data

```{r boundary_hz, eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
# hz data
mic <- read.table(file='../data/HZ/hz_asv_2018.txt',
                  header=TRUE,row.names=1) # 51441 taxa in total
# construct column id
bay_loci <- c('HB1.1','HB1.2','HB1.3','HB2.1','HB2.2','HB2.3',
              'HB3.1','HB3.2','HB3.3','HB4.1','HB4.2','HB4.3',
              'HB5.1','HB5.2','HB5.3','HB6.1','HB6.2','HB6.3',
              'HB7.1','HB7.2','HB7.3','HB8.1','HB8.2','HB8.3',
              'HB9.1','HB9.2','HB9.3','HB10.1','HB10.2','HB10.3')
era_sy_loci <- c('SY1.1','SY1.2','SY1.3','SY2.1','SY2.2','SY2.3',
                 'SY3.1','SY3.2','SY3.3','SY4.1','SY4.2','SY4.3',
                 'SY5.1','SY5.2','SY5.3','SY6.1','SY6.2','SY6.3')
era_jx_loci <- c('JX1.1','JX1.2','JX1.3','JX2.1','JX2.2','JX2.3',
                 'JX3.1','JX3.2','JX3.3','JX4.1','JX4.2','JX4.3',
                 'JX5.1','JX5.2','JX5.3','JX6.1','JX6.2','JX6.3')

# for hz bay
mic_bay <- mic[,bay_loci] # 51441 taxa
# filter records with zero counts across all sample sites
mic_bay <- mic_bay[rowSums(mic_bay)>0,] # 24383 taxa
load('../output/HZ/param_trio_bay.RData')
id_vag_lag <- classify_vag_lag(mic_bay, param_trio)
save(id_vag_lag, file='../output/HZ/id_vag_lag_bay.RData')
# hz era
# for hz era
mic_era <- mic[,c(era_sy_loci,era_jx_loci)] # 51441 taxa
# filter records with zero counts across all sample sites
mic_era <- mic_era[rowSums(mic_era)>0,] # 33606 taxa
load('../output/HZ/param_trio_era.RData')
id_vag_lag <- classify_vag_lag(mic_era, param_trio)
save(id_vag_lag, file='../output/HZ/id_vag_lag_era.RData')
```

### AO data

```{r boundary_ao, eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
# ao data
# import the whole data set
mic <- read.table('../data/AO/ao_asv.txt')
# industrial factory names
id_dye <- c('LS','SF','CZ','BA','YF')
id_med <- c('GB','YTSW','XHC','ZC')
id_pes <- c('YT','YN')
lis_ind <- list(id_dye, id_med, id_pes)
names(lis_ind) <- c('Dye','Medicine','Pesticide')
# ao procedures id
id_inf <- c('Inf1','Inf2','Inf3')
id_axi <- c('Ax1','Ax2','Ax3')
id_oxi <- c('Ox1','Ox2','Ox3')
id_eff <- c('Eff1','Eff2','Eff3')
lis_pro <- list(id_inf, id_axi, id_oxi, id_eff)
names(lis_pro) <- c('influent','anoxic','oxic','effluent')

# ao influent
for(region in 1:4){
  id_col <- c()
  for(i in 1:length(lis_ind)){
    id_ind <- lis_ind[[i]]
    id_pro <- lis_pro[[region]]
    id_col <- c(id_col,
                paste0(rep(id_ind, each=length(id_pro)),'_',rep(id_pro, length(id_ind))))
  }
  mic_sub <- mic[,id_col]
  mic_sub <- mic_sub[rowSums(mic_sub)>0,]
  
  load(paste0('../output/AO/param_trio_',names(lis_pro)[region],'.RData'))
  id_vag_lag <- classify_vag_lag(mic_sub, param_trio)
  save(id_vag_lag, file=paste0('../output/AO/id_vag_lag_',names(lis_pro)[region],'.RData'))
  
}
```

## Visualize the community structure as the admixture of dispersal vanguards and laggards

### HZ data

#### HZ Bay

```{r vag_lag_hz_bay, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/HZ/param_trio_bay.RData')
load('../output/HZ/id_vag_lag_bay.RData')
plot_vag_lag(param_trio, id_vag_lag, zoom_in=TRUE)
```

#### HZ ERA

```{r vag_lag_hz_era, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/HZ/param_trio_era.RData')
load('../output/HZ/id_vag_lag_era.RData')
plot_vag_lag(param_trio, id_vag_lag, zoom_in=TRUE)
```

### AO data

#### AO *Influent*

```{r vag_lag_ao_influent, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/AO/param_trio_influent.RData')
load('../output/AO/id_vag_lag_influent.RData')
plot_vag_lag(param_trio, id_vag_lag, zoom_in=TRUE)
```

#### AO *Anoxic*

```{r vag_lag_ao_anoxic, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/AO/param_trio_anoxic.RData')
load('../output/AO/id_vag_lag_anoxic.RData')
plot_vag_lag(param_trio, id_vag_lag, zoom_in=TRUE)
```

#### AO *Oxic*

```{r vag_lag_ao_oxic, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/AO/param_trio_oxic.RData')
load('../output/AO/id_vag_lag_oxic.RData')
plot_vag_lag(param_trio, id_vag_lag, zoom_in=TRUE)
```

#### AO *Effluent*

```{r vag_lag_ao_effluent, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
load('../output/AO/param_trio_effluent.RData')
load('../output/AO/id_vag_lag_effluent.RData')
```


```{r SI, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
sessionInfo()
```

